---
title: "Genres of Music"
subtitle: "b565 Bonus Homework"
author: "Kevin O'Neill"
date: "12/18/2015"
output: pdf_document
---

Question 1:
The relative frequencies of all 6 genres are as follows:
First we find the total number and then find the relative frequency by dividing the frequency of each
Genre by the total.
```{r}
g <- read.csv("genresRaw.csv", header=TRUE)

g.nrows <- nrow(g)
g.splits <- split(g, g$GENRE)
g.names <- names(g.splits)
freqs <- c(nrow(g.splits[['Blues']]),
           nrow(g.splits[['Classical']]),
           nrow(g.splits[['Jazz']]),
           nrow(g.splits[['Metal']]),
           nrow(g.splits[['Pop']]),
           nrow(g.splits[['Rock']]))

dim(g)

g.freq <- data.frame(Genre = g.names, Freq = freqs, Rel_Freq = freqs/g.nrows)
g.freq

# A quicker way to do this 
round(prop.table(table(g$GENRE)) * 100, digits = 2)
```


****
Question 2: We can look at the pairwise correlations of all the features by doing the following:
```{r}
require(Hmisc)
g.sub <- g[c(-1:-4)] # reduce data.frame to features only
g.mat <- as.matrix(g.sub) # convert to matrix
g.cor <- rcorr(g.mat, type="pearson")
# write.csv(g.cor$r, "genre-correlation.csv")
g.cor.mat <- g.cor$r 
# delete extra row and column that produced the NAs
g.cor.mat <- g.cor.mat[,-192]
g.cor.mat <- g.cor.mat[-192,]
dim(g.cor.mat)
g.cor.mat[1:5, 1:5]
```


****
Question 3:
We can find the number of performers by checking for the unique ARTIST_ID's
```{r}
length( unique(g$ARTIST_ID))
```
There are 60 performers present in the data.


****
Question 4: 

```{r}
artists <- unique(g$ARTIST_ID)
num_pieces <- c()
for (artist in artists) {
    artist.df <- g[g$ARTIST_ID == artist, ]
    num_pieces <- append(num_pieces, length(unique(artist.df$PIECE_ID)))
}
num_pieces_per_artist <- data.frame(ARTIST_ID = artists, NUM_PIECES = num_pieces)
# These are the counts for the number of pieces each performer contributed
num_pieces_per_artist
```


****
Question 5:

Every piece is subdivided into 21 segments

****
Question 6:

```{r}
require(class)
require(gmodels)
# using knn algorithm to classify genres
g.train <- read.csv("genresTrain.csv", header=TRUE)
g.test <- read.csv("genresTest.csv", header=TRUE)
g.train[1:5, 1:10]
# function to normalize the columns
normalize <- function(x) {return ((x - min(x)) / (max(x) - min(x)))}
g.n <- as.data.frame(lapply(g.train[1:191], normalize)) # normalizing numeric columns
g.n.train <- g.n[1:6247,]
g.n.test <- g.n[6248:12495,]
#g.n.test <- as.data.frame(lapply(g.test[1:191], normalize))
#g.n.labels <- g.train$GENRE # vector of GENRE labels
g.n.train.labels <- g.train$GENRE[1:6247]
g.n.test.labels <- g.train$GENRE[6248:12495]

k <- round(sqrt(nrow(g.n.train)))
genre_test_pred <- knn(g.n.train, g.n.test, g.n.train.labels, k)
CrossTable(x = g.n.test.labels, y = genre_test_pred, prop.chisq=FALSE)
```


****
Question 7:
```{r}


```
